A game-theory modeling approach to utility and strength of interactions dynamics in biomedical research social networks

### 2017-04-04 casm.tex
these instead of this
Included new references and also the discussion of results considering the references at the itroduction.

%%% Load packages
%\usepackage{amsthm,amsmath}
%\RequirePackage{natbib}
%\RequirePackage[authoryear]{natbib}% uncomment this for author-year bibliography
%\RequirePackage{hyperref}
\usepackage[utf8]{inputenc} %unicode support
\usepackage{color}
% \usepackage[applemac]{inputenc} %applemac support if unicode package fails
%\usepackage[latin1]{inputenc} %UNIX support if unicode package fails

@@

{\color{red}In a similar vein to the work of Li and Cao,} we explore the network
effect on the distribution of players having access to certain amount
of resources shared by other players in the network and the
distribution of the strength of interactions among them. Particularly,
we implemented two games played simultaneously: one for maximizing
individual utility based on the iterated Prisoner's Dilemma; the
other, a coordination game for maximizing the connection strength
between players. We are interested in how they affect each other in
the context of a network of scientific collaboration under the idea
that while researchers are interested in maximizing their utilities,
they also know that it is important to invest in building
collaborative relationships. These two behaviors are explored in a
biomedical research community in M\'exico.\\

@@

@@

  There are two issues we consider important. The first one is about
  confronting our results with data available to us from the biomedical research
  community. Second, we would like to comment on our future work.\\

  The two main results of our model are the smoothness in the phase
  transition-like behavior for different parameters and the distribution of
  utility and strength of interactions in the FOSISS network compared to other
  networks and their topologies. We are certain that the particular structure of
  FOSISS network is playing a central role in the results and because of that we
  would like to discuss it a little bit further. FOSISS network is highly
  hierarchical according to its heterogeneity of $0.873$. Because of it, one
  would expect to find the presence of important hubs \cite{Wu:2008}, that is, a
  few researchers control the whole network, as has been reported in some other
  places \cite{Yousefi-etal:2008}. Surprisingly, FOSISS network centralization
  is very low $0.023$, i.e. there are no researchers that centralize the
  majority of connections. Our guess is that the network is composed of many
  small communities or groups with a central researcher or Principal
  Investigator (PI). If this is the case in the FOSISS network, it means that
  those groups have a very hierarchical structure as well.\\

  Under such structure, when the clustering coefficient is
  considered, it can be said that groups are also well connected but
  that inter-group connections are sparse. In other words, individual
  groups are strongly connected within but the network as a whole is
  supported by a small number of links. We came by this idea partially
  from another study about scientific collaborations based on
  co-authorships in one of the research centers that is part of the
  FOSISS network \cite{HernandezLemus:2013}. In the cited reference, an
  apparently well integrated community was found (high clustering
  coefficient and a very short characteristic path length). Such
  integration was mostly superficial, since it depended on the
  presence of external collaborators from other research institutions
  (most of them from overseas).  Removing these external collaborators
  brakes down the network into small subgraphs that worked
  independently. Remarkably, those subgraphs corresponded to the real
  groups of that research center.  What is more, several groups had a
  hierarchical structure as the one we suspect is common in FOSISS
  subgraphs. The results showed that collaboration was poor between
  groups but strong among the members of each group, and that
  collaboration among groups doesn't emerge bottom-up, instead it
  seems to be promoted from the top, from the administrative
  authorities.\\

@@

 As is common among scientific communities in biomedical
  research, PIs play a central role in the network. Strong PIs and
  well connected groups seem to be somehow responsible for the high
  levels of strength of interactions and the centralization of utility
  in our simulation. As for what seems to be phase transitions, in the
  case of FOSISS networks, these are smoother than those in the other
  networks with different topologies, even for those with a
  small-world topology. We think that this behavior is also the result
  of the hierarchical structure already mentioned. If this is true,
  strength of interactions is first lost in the edges that link
  different groups and then in the edges that connect members of the
  groups. Connections between groups would not be as dense as those
  inside the groups, which means that there would not be enough
  information on the behavior of one group regarding its neighbors to
  constrain them as it seems to happen with individual researchers
  inside their communities. Nevertheless, if values of the parameter
  of cooperation $\nu$ keep increasing and it becomes more difficult to
  strengthen interactions, then strength of interactions begins to
  diminish inside groups.\\

>>>>

%%% Load packages
%\usepackage{amsthm,amsmath}
%\RequirePackage{natbib}
%\RequirePackage[authoryear]{natbib}% uncomment this for author-year bibliography
%\RequirePackage{hyperref}
\usepackage[utf8]{inputenc} %unicode support
\usepackage{color}
\usepackage{cite}
% \usepackage[applemac]{inputenc} %applemac support if unicode package fails
%\usepackage[latin1]{inputenc} %UNIX support if unicode package fails

@@

{\color{red}In a similar vein to the work of Li and Cao,} we explore the network
effect on the distribution of players having access to certain amount of
resources shared by other players in the network and the distribution of the
strength of interactions among them. Particularly, we implemented two games
played simultaneously: one for maximizing individual utility based on the
iterated Prisoner's Dilemma (PD); the other, a coordination game for maximizing the
connection strength between players. We are interested in how they affect each
other in the context of a network of scientific collaboration under the idea
that while researchers are interested in maximizing their utilities, they also
know that it is important to invest in building collaborative
relationships. \\ %This is important because what we have seen

@@

{\color {red}From our anthropological fieldwork among biomedical research communities, their
networks appear to be quite paradoxical: while they seem very
hierarchical and only few researchers get the most of the profit of collective
work, a great deal of collaboration is maintained. In other words, we are
intrigued by the fact that for all the power and profit gain asymmetries,
people are still willing to collaborate, build trust and remain loyal to their
PI's and mentors.These two behaviors are explored in a biomedical research
community in M\'exico. We believe that by implementing PD game and coordination
game in a real world network is a way of capturing the essence of this
problem. From our point of view this is an important contribution as we not only
use data extracted from bibliometric source as most collaboration networks are
studied, but that the founding question on which the model has been developed is
empirically based in the sociology of the case.}\\

@@

% There are several issues we consider important to discuss. The first one is about confronting
% our results with data available to us from the biomedical research
% community. Second, we would like to comment on our future work.\\

{\color {red}In the area of complex networks and game theory, it has being affirmed for a
long time that scale-free networks promote the evolution of cooperation. Part of
this property has been attributed to the presence of hubs
\cite{Santos-etal:2005,Szolnoki-etal:2008,Du-etal:2008,Li-etal:2009,Yang-etal:2009,Lei-etal:2009,Liu-etal:2010,Cheng-etal:2010B,Li-etal:2014,Li-etal:2015,Ichinose:2017}. Perhaps
less attention has been place on the persistence of cooperation and the the role
played by the clustering coefficiente in scale-free and small-world networks
\cite{Kim-etal:2002,Masuda-etal:2003,Wu-etal:2005,Wu-etal:2006,Guan-etal:2006,Fu:2007,Zhang-etal:2007,Thibert-Plante-etal:2007,Chen-etal:2008}. This is where our result comes in.} \\

The two main outcomes of our model are the smoothness in the phase
transition-like behavior for different parameters and the distribution of
utility and strength of interactions in the FOSISS network compared to other
networks and their topologies. We are certain that the particular structure of
FOSISS network is playing a central role in the results and because of that we
would like to discuss it a little bit further. FOSISS network is highly
hierarchical according to its heterogeneity of $0.873$. Because of it, one would
expect to find the presence of important hubs \cite{Wu:2008}, that is, a few
researchers control the whole network, as has been reported in some elsewhere
\cite{Yousefi-etal:2008}. Surprisingly, FOSISS network centralization is very
low $0.023$, i.e. there are no researchers that centralize the majority of
connections. Our guess is that the network is composed of many small communities
or groups with a central researcher or Principal Investigator (PI). If this is
the case, it means that those groups have a very hierarchical structure as
well. This is similar to what has been reported by others about the
heterogeneity in small-world networks, in which the combination of certain
amount of hierarchy and small-worldness makes cooperation to peak over compared
to the evolution of cooperation in homogeneous and the most heterogeneous
networks \cite{Fu:2007}. For the FOSISS netwok, this property would enhanced
cooperation among players, eventhough the distribution of utilities are skewed
towards a handful of researchers. \\

% The two main results of our model are the smoothness in the phase
% transition-like behavior for different parameters and the distribution of
% utility and strength of interactions in the FOSISS network compared to other
% networks and their topologies. We are certain that the particular structure of
% FOSISS network is playing a central role in the results and because of that we
% would like to discuss it a little bit further. FOSISS network is highly
% hierarchical according to its heterogeneity of $0.873$. Because of it, one would
% expect to find the presence of important hubs \cite{Wu:2008}, that is, a few
% researchers control the whole network, as has been reported in some other places
% \cite{Yousefi-etal:2008}. Surprisingly, FOSISS network centralization is very
% low $0.023$, i.e. there are no researchers that centralize the majority of
% connections. Our guess is that the network is composed of many small communities
% or groups with a central researcher or Principal Investigator (PI). If this is
% the case in the FOSISS network, it means that
% those groups have a very hierarchical structure as well.\\

Under such structure, when the clustering coefficient is considered, it can be
said that groups are also well connected but that inter-group connections are
sparse. In other words, individual groups are strongly connected within but the
network as a whole is supported by a small number of links. {\color{red}This is
  not an uncommon pattern in social networks. It has been claimed that this form
  of structure has been evolutionarily advantageous
  \cite{Turchin:2015}. Moreover, research groups internal cohesion may be
related to Wardil and Hauert idea that collaboration among scientists --at least
in the case of multi-authored articles-, is more similar to a ``snow-drift
game'' than to a PD game, since in the former everybody wins or everybody pays a
cost but the cost are not evenly distributed among the players.}\\

We came by the idea that intra-networks are stronger than inter-networks, from
another study about scientific collaborations based on co-authorships 
\cite{HernandezLemus:2013}. In the cited reference, an apparently well
integrated community was found (high clustering coefficient and a very short
characteristic path length). Such integration was mostly superficial, since it
depended on the presence of external collaborators from other research
institutions (most of them from overseas).  Removing these external
collaborators brakes down the network into small subgraphs that worked
independently. Remarkably, those subgraphs corresponded to the real groups of
that research center.  What is more, several groups had a hierarchical structure
as the one we suspect is common in FOSISS subgraphs. The results showed that
collaboration was poor between groups but strong among the members of each
group, and that collaboration among groups doesn't emerge bottom-up, instead it
seems to be promoted from the top, from the administrative
authorities.\\

@@

As is common among scientific communities in biomedical research, PIs play a
central role in the network. Strong PIs and well connected groups seem to be
somehow responsible for the high levels of strength of interactions and the
centralization of utility in our simulation. As for what seems to be phase
transitions, in the case of FOSISS networks, these are smoother than those in
the other networks with different topologies, even for those with a small-world
topology. We think that this behavior is also the result of the hierarchical
structure already mentioned. If this is true, strength of interactions is first
lost in the edges that link different groups and then in the edges that connect
members of the groups.\\

Connections between groups would not be as dense as those
inside the groups, which means that there would not be enough information on the
behavior of one group regarding its neighbors to constrain them as it seems to
happen with individual researchers inside their communities. Nevertheless, if
values of the parameter of cooperation $\nu$ keep increasing and it becomes more
difficult to strengthen interactions, then strength of interactions begins to
diminish inside groups. {\color {red}And yet, the smoothness observerd in the small-world
network and FOSISS network may be attributed to the power of the clustering
coefficient. As has been referred by Li and Cao \cite{Li-etal:2009}, and
Thibert-Plante and Parrott \cite{Thibert-Plante-etal:2007} for scientific
collaboration networks, high clustering coefficient allows the system to resist
the invasion of non-cooperators.}\\
